Image Caption Generation With CLIP+GPT-2 Model

### 2025-05-04 Latex_Paper/Paper.tex
update all

  Our model efficiently manages unlabeled images, reducing the need for manual effort.  
  \item Effective CLIP-to-GPT-2 Mapping:

  Our multi-head attention mechanism bridges visual and textual modalities for high-quality captions.
  \item Noise Injection Boosts Generalization:

  Introducing noise during training boosts generalization and helps create diverse, accurate captions.
\end{itemize}

\subsection{Future Work}
\begin{itemize}
  \item 
  \item 
  \item 
\end{itemize}

\section{Conclusion}

>>>>

    Our model demonstrates strong automatic organization 
    capabilities for visual content, especially when dealing 
    with a large number of unlabeled images. By generating 
    relevant and informative captions, the system can group 
    images without manual annotation.

        \item Effective CLIP-to-GPT-2 Mapping:

    The key design of our entire model is the multi-head 
    attention mechanism used to connect CLIP visual embedding
    and GPT-2 language model input. Unlike simple linear projection, 
    the multi-head attention mechanism enables the model to learn 
    multiple semantic alignment patterns between image features and 
    different dimensions of generated captions. Test results show 
    that the implementation of this mechanism can generate more 
    accurate captions that reflect the details of image semantics.

    \item Noise Injection Boosts Generalization:
    We observe that adding small Gaussian noise into CLIP embeddings 
    during training improves the model's generalization across 
    different image inputs. This regularization technique helps 
    the mapping module avoid overfitting to specific embedding 
    patterns and encourages the model to generate captions that
    are both accurate and diverse. Result shows that noise
    improves both BLEU and CIDEr scores, indicating improved
    robustness and semantic coverage of caption generation.

\subsection{Future Work}
In future work, we plan to expand the test 
dataset to cover more diverse and challenging images to 
more robustly evaluate the generalization ability of our model. 
In addition, we can try to study and combine more powerful visual 
language models (such as CLIP ViT-L or GPT-3 variants) to achieve 
stronger semantic alignment capabilities.

\section{Conclusion}
In this project, we implemented an image captioning 
system by combining the powerful visual encoding capabilities 
of CLIP and the generative language capabilities of GPT-2. By 
introducing a mapping module based on a multi-head attention 
mechanism, we effectively reduced the gap between visual and 
textual modalities, enabling the generation of fluent and 
semantically relevant captions. We further enhanced the 
generalization ability of the model by adding noise during training, and selected appropriate loss functions and objective functions for training. Finally, evaluations using BLEU and CIDEr metrics confirmed the effectiveness of our model, especially in capturing human-written semantics, with the CIDEr score reflecting its excellent performance.
